---
title: Allocating MSOAs to a hexmap grid
date:  January 17, 2021
output: 
  html_document:
    keep_md: TRUE
---



```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE)

# libraries ---------------------------------------------------------------
{
  library(here)
  library(dplyr)
  library(jogger)
  library(purrr)
  library(sf)
  library(tmap)
  library(wesanderson)
  library(ggsci)
}

# tmap -------------------------------------------------------------
tmap_mode("plot")

tmap_options()


# mapping fill palette ---------------------------------------------

# my_pal <- ggsci::pal_futurama()
my_pal <- ggsci::pal_d3("category20")

```


### LADs, MSOAs and Centroids

The intention is to create a [hexmap](hexmap) of MSOAs in the wider county of Sussex.
This area comprises three modern-day upper-tier local authorities:

* Brighton and Hove
* West Sussex
* East Sussex.

First, we obtain the geographical data we will need to use:


```{r get-bounds, echo = TRUE, eval = FALSE}
# Local Authority District names and boundaries within 3 upper-tier authorities
sussex_lads <- c("Brighton and Hove", "East Sussex", "West Sussex") %>%
  purrr::map_df(
    ~ jogger::geo_get("lad"
                      , .
                      , "utla"
                      , return_style = "simple"
                      , spatial_ref = 27700
                      )
    ) %>% 
  dplyr::rename(lad20nm = ltla20nm)


# MSOA names and boundaries within each LAD
sussex_bounds <- sussex_lads %>%
  dplyr::pull(lad20nm) %>%
  purrr::map_df(
    ~ jogger::geo_get("msoa"
                      , .
                      , "lad"
                      , spatial_ref = 27700
                      , shape_fields = TRUE
                      )
    ) %>% 
  dplyr::relocate(c(shape_area, shape_length, geometry), .after = last_col())
  # # not needed now - was previous way of sorting areas by smallest MSOA area
  # %>% order_along(lad20cd, shape_area)


# Population-weighted centroids of the MSOAs
sussex_centroids <- sussex_bounds %>%
  sf::st_drop_geometry() %>% 
  dplyr::select(starts_with(c("msoa", "lad"))) %>% 
  dplyr::left_join(
    pull(., msoa11cd) %>% 
    jogger::geo_get_bounds(
      "msoa11cd"
      , .
      , spatial_ref = 27700
      , return_centroids = TRUE
    )
  ) %>% 
  sf::st_as_sf()


```




```{r load-bounds, include = FALSE}

# myrmidon::save_it(sussex_lads)
# myrmidon::save_it(sussex_bounds)
# myrmidon::save_it(sussex_centroids)

sussex_lads <- readRDS(here::here("rds_data", "sussex_lads.Rds"))
sussex_bounds <- readRDS(here::here("rds_data", "sussex_bounds.Rds"))
sussex_centroids <- readRDS(here::here("rds_data", "sussex_centroids.Rds"))

```



West Sussex and East Sussex then comprise, between them, 12 lower-tier authorities (LADs), making 13 in total (Brighton and Hove is a unitary.)

The area contains `r nrow(sussex_bounds)` Middle Layer Super Output Areas (MSOAs).
These are Census geographies (divisions) that are designed to have populations of roughly 8,000 people each, though this varies quite a bit in reality
(and due to change since their establishment in 2011).

MSOAs are each located within a certain LAD - they do not cross LAD boundaries.

The [ONS](ons-opengeo) provides population-weighted centroids (central points) for each MSOA.


Now we have data for the boundaries of the LADs:

```{r lads-list}
head(sussex_lads, 3)


```

and we can plot these along with their (geometric) centroids:

```{r lads-list2}

sussex_lad_centroids <- sussex_lads %>% 
  sf::st_centroid()

tm_shape(sussex_lads) +
  tm_borders("grey45", lwd = 2) +
  tm_fill("lad20nm", alpha = 0.2, palette = my_pal(13)) +
  tm_layout(legend.outside = TRUE) +
  tm_shape(sussex_lad_centroids) +
  tm_dots(col = "orange", size = 0.1)

```


We also have the boundaries of the MSOAs:

```{r msoas-list}

head(sussex_bounds, 3)



```


which we can plot as well as their centroids:

```{r cent-list}

tm_shape(sussex_bounds) + 
  tm_borders("grey75", lwd = 1) + 
  tm_shape(sussex_centroids) + 
  tm_dots(col = "white", size = 0.2) + 
  tm_shape(sussex_centroids) + 
  tm_dots(col = "tomato2", size = 0.1) 


```


### Using `sf` to make a hex grid for our area









[hexmap]: https://github.com/houseofcommonslibrary/uk-hex-cartograms-noncontiguous#why-use-a-non-contiguous-cartogram-and-what-about-other-options
[ons-opengeo]: https://
