---
title: "working out the algorithm"
vignette: >
  %\VignetteIndexEntry{algorithm}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(msoahexmap)
```

## The intention

The intention is to make the best hexmap of the MSOAs in England (and Wales
  maybe) that I can.

I'm hoping I can learn something about working with spatial functions and
  designing a complex algorithm to achieve what I want.

I want the process to be as automated as possible, for two reasons:
* I want it to be reproducible and relatively easy to improve iteratively
* I want to have the core of an algorithm that is portable to other situations
  and projects. If there is too much manual intervention required to get the map
  right, the process isn't really portable or generalisable.

## The problem part 1

There are lots of different ways you could try to do this.

One attempt that was very popular on twitter during the Covid-19 years was
  developed by Carl Baker and colleagues at the House of Commons Library.


## The problem part 2: London

Yeah, London. That London.

You're a problem. A big problem.

We'll deal with you later.






## London

You know earlier I said I would deal with London later?

That was mainly for simplicity in following some kind of logical order in
  writing this.
  But in truth I had actually dealt with London first.

The problem is population density.

If you make your hexagons small enough to fit all 931 of London's into
  the actual geographical shape and size of London on your grid, most of the
  rest of England (let alone Wales) has hexes so sparse that they look
  ridiculously out of proportion in the vast rural swathes of much of
  England.
  There are exceptions of course, areas of relatively high density in the
  regional cities.
  But it's only the North West that comes anywhere close to having the kind of
  population density, in the Manchester-Liverpool metropolitan region, that
  London does.

And the North West still has Cumbria to even things out.
  Needless to say, Bromley and Havering - don't even think about it: you aren't
  going to compete with the Lake District.

So what to do with London?

My solution is simple: throw it in the North Sea.


### Processing London

In order to get all of London's MSOAs represented as well as possible within
  its actual shape, so that we can give the rest of the country hexes as big as
  possible, I actually had to be quite interventionist about what went where,
  and in what order the districts of London got allocated their hexes.
  I took the liberty however of removing certain hexes from the grid to
  represent the lower Thames, the lower Lea, and Richmond Park, and some other
  bits around the fringes of the city.

My original work on London used a hex size just small enough to fit everything
  in and to allow for the bits of shaping just mentioned.
  But otherwise it was as big as possible.

London original hexagon size: xxxx

But still smaller than we would ideally like to use for the rest of the country.
So, leaving London aside, I moved onto the next densest region in England: the
  North West.
  Again using the largest possible hex size that allowed me to fit everything
  into the actual shape and size of the region without areas being wildly out
  of place or distorted, I began processing the North West.
  Again this was quite interventionist.
  But once this region is done, the other regions should fall into place
  relatively easily, with their generally lower densities.
  (West and South Yorkshire were also a bit of a challenge, mainly because they
  adjoin much of the North West region so there is not much scope for moving
  them around.)

### Hex size: it's all about compromise


## Attempt 1

It's a little sad to say that I originally started working on this in about
  2021, made some good progress but never finished it off, then came back to it
  intermittently through 2022 but it's now August 2024 and this still hasn't got
  finished off.

I have been tinkering.

My first attempts were region-based. 